---
title: "Homework 07"
author: "Christina Lee"
date: "10/26/2021"
output: pdf_document
---



Is there an effect of Age on IQ? Please perform all calculations by hand using the equations in the lessons unless otherwise specified.


1. Plot these four points using R.
```{r}
# first create the table and save it as data frame. 
ageIQ <- data.frame(Age = c(23,18,10,45), IQ = c(100,105,95,120))
ageIQ

# create the plot. 
library(ggplot2)
ggplot(data=data.frame(ageIQ), mapping=aes(x=Age, y=IQ)) + geom_point() + ggtitle("Age V.S IQ") +
  theme(plot.title= element_text(hjust=0.5))



```






2. Calculate the covariance between age and IQ.

$$cov(x,y) = \frac{1}{n-1} \sum_i{(x_i-\bar{x})(y_i-\bar{y})}   $$
$$cov(x,y) = \frac{(23-24)(100-105)+(18-24)(105-105)+(10-24)(95-105)+(45-24)(120-105)}{3}$$
$$ cov(x,y) = 153.\bar{3}   $$

```{r}
x <- c(23,18,10,45)  # Age 
y <- c(100,105,95,120) # IQ

# covariance calculated by R to verify hand calculation.
cov(x,y)



```









3. Calculate their correlation. What does the number you get indicate?

Sx = 14.98 \newline
Sy = 10.80 \newline
$$ r= \frac{cov(x,y)} {S_xS_y}     $$


$$ r = \frac{153.3}{(14.98)(10.80)}     $$
$$ r = 0.95   $$

```{r}
x <- c(23,18,10,45)  # Age 
y <- c(100,105,95,120) # IQ

# correlation calculated by R to verify hand calculation.
cor(x,y)


```
The correlation calculated by hand is 0.95 which is fairly close to 1. This indicates that when x (Age) is higher than its mean, y (IQ) is as well -- Age and IQ are positively correlated. \newline
\newline










4. Calculate the regression coefficients B0 and B1 and write out the equation of the best-fit line relating
age and IQ.

First solve for B1:
$$  r = \frac{cov(x,y)}{S_xS_y} = \beta_1 \frac{S_x}{S_y} $$
$$ \frac{r}{\frac{S_x}{S_y}} = \beta_1 $$
$$ \beta_1 = \frac{0.95}{\frac{14.98}{10.80}}  $$
$$ \beta_1 = 0.68  $$


Once we solved for B1, we can now solve for B0: 

$$ \beta_0 = \hat{y} - \beta_1 \hat{x}  $$

$$ \beta_0 = 105 - 0.68 * 24     $$
$$ \beta_0 = 88.68   $$
Therefore, the best-fit line equation is then: 


$$ y= 88.68 + 0.68 x  $$ 


5. Calculate the predicted yi for each xi.

$$ 104.32= 88.68 + 0.68 (23)  $$
$$ \hat{y_i} = 104.32$$

$$100.92= 88.68 + 0.68 (18)  $$
$$ \hat{y_i} = 100.92$$



$$95.48= 88.68 + 0.68 (10)  $$
$$ \hat{y_i} = 95.48$$



$$119.28= 88.68 + 0.68 (45)  $$

$$ \hat{y_i} = 119.28$$



6. Calculate Rˆ2 from the TSS/SSE equation. How does it relate to the correlation? What does the number you get indicate?



$$ r^2 = \frac{TSS-SSE}{TSS}$$
$$ TSS = \sum_i({y_i - \bar{y}})^2 $$
$$ TSS = \sum_i ({100-105})^2 +  ({105-105})^2 +({95-105})^2 + ({120-105})^2  $$
$$ TSS = 350 $$ 

$$ SSE = \sum_i(y_i - \hat{y_i})     $$

$$ SSE = \sum_i (100-104.32)^2 + (105-100.92)^2 + (95 - 95.48)^2  + (120-119.28)^2  $$
$$ SSE = 36.05   $$
$$ r^2 = \frac{350-36.05}{350}   $$
$$ r^2 = 0.897 $$ 


r^2 is a measure of the proportion of the total variation in y explained by x --how strong/weak are the variables correlated with each other. r^2 is also sometimes known as the proportional reduction in error so, we can then say we have reduced the error by 89.7%, or 89.7% of the variation in y is explained by the model --x and y are very much correlated. 






7. Calculate the standard error of B1, and use that to test (using the t test) whether B1 is significant.


Let's first calculate the se for y hat. 

$$ se_{\hat{y}}  = \sqrt {\frac{\sum({y_i-\hat{y}})^2}{n-2} }   $$
  

$$ se_{\hat{y}} = \sqrt{\frac{SSE}{n-2}}  $$


$$se_{\hat{y}} = \sqrt{\frac{36.05}{4-2}}    $$



  
$$ se_{\hat{y}} = 4.24  $$

Once we've calculated the se for y hat, solve for se of B1: 

$$se_{\beta_1} =se_{\hat{y}} \sqrt{\frac{1}{\sum{(x_i - \bar{x})^2}}}$$   

Solve for se of B1: 

$$ se_{\beta_1} = 4.24 \frac{1}{\sqrt{674}}     $$
$$ se_{\beta_1} = 0.16$$



Now, conduct a t test to see whether B1 is significant: 

$$ H_a: \beta_1  \ne 0   $$

$$ H_o: \beta_1  = 0   $$

$$ Test\:Statistic = \frac{\beta_1 - \mu_0}{se_{\beta_1}}    $$
$$ Test\: Statistics = \frac{0.68 - 0} {0.16}$$

$$ Test \: Statistics = 4.25   $$ 

$$ df = n - k - 1  $$

$$ df = 4 - 1 - 1 $$ 


$$ df = 2 $$ 

Now, calculate the threshold value for two-tailed test:
```{r}
qt(0.975,2) # upper threshold
qt(0.025,2) # lower threshold

```


$$ T_{critical} = \pm4.30$$
Results: \newline
 Since our t statistic value = 4.25 is not greater than 4.30, nor less than -4.30 the threshold value we fail to reject the null. This indicates that the positive correlation between age and IQ is not statistically significant --the change in age does not really produce a change in IQ and is just a statistical illusion. \newline
 




8. Calculate the p-value for B1 and interpret it.


```{r}
2*pt(4.25,2,lower.tail=F)


```

Since p-value 0.05115253 is > 0.05 we again fail to reject the null hypothesis --the effect of age on IQ is not statistically significant.\newline









9. Calculate the 95% CI for B1 and interpret it.


$$ 95 \% \:CI = \beta_1 = 0.68 \pm 4.30 *0.16 = [-0.008,1.368]  $$

We are 95% certain that as age increases per year, it will result in an increase in IQ between -0.008 and 1.368, with our best guess being 0.68 increase in IQ as age increases each year. \newline
\newline







10. Confirm your results by regressing IQ on Age using R

Results do match hand calculations above. 
```{r}
biv_model <- (lm(IQ ~Age, data =ageIQ)) 
summary(biv_model)


```










11. Plot your points again using R, including the linear fit line with its standard error.





```{r}

# create the plot with best-fit line and standard error. 
library(ggplot2)
ggplot(data=data.frame(ageIQ), mapping=aes(x=Age, y=IQ)) + geom_point() + 
  geom_smooth(formula = y ~ x, method = "lm") + ggtitle("Age V.S IQ") +
  theme(plot.title= element_text(hjust=0.5))




```












12. What are you final conclusions about the relationship between age and IQ?


According to our results, r = 0.95 shows that age and IQ are positively correlated. With r^2 = 0.897, it suggests that we have reduced the error by 89.7% and 89.7% of the variation in y can be explained by the model. However, the t-test that we have conducted and with the p-value and t-statistics tells us that the positive correlation between age and IQ that we have witnessed is not, in fact, statistically significant --the change in age does not really produce a change in IQ and it is just a statistical illusion.





















